High Performance Computing for Global Optimization Problems

### 2017-11-03 2018/ITNT/ITNT2018.tex
Pictures ans tables

\Russian
Параллельный индексный алгоритм, который может быть применен для решения подобных задач с частично вычислимыми функциями и который реализован в системе Globalizer, подробно описан в [2]. Основная идея алгоритма состоит в сведении исходной многомерной задачи к набору связанных подзадач меньшей размерности, и параллельному их решению. Схемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.

>>>>

\Russian
ѕараллельный индексный алгоритм, который может быть применен дл€ решени€ подобных задач с частично вычислимыми функци€ми и который реализован в системе Globalizer, подробно описан в [2]. ќсновна€ иде€ алгоритма состоит в сведении исходной многомерной задачи к набору св€занных подзадач меньшей размерности, и параллельному их решению. —хемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.

\section{Dimensionality reduction}

\subsection{Dimensionality reduction using space-filling curves}

The use of Peano curve $y(x)$ 
\begin{equation}
\left\{y\in R^N: -2^{-1}\leq y_i \leq 2^{-1}, 1 \leq i \leq N\right\}=\left\{y(x):0\leq x \leq 1 \right\}
\end{equation}
unambiguously mapping the interval of real axis $[0,1]$ onto a $N$-dimensional cube is the first of the dimensionality reduction methods considered.  To implement this method of dimensionality reduction a numerically constructed curve (\textit{evolvent}) is used. The evolvent is $2^{-m}$ accurate approximation of the theoretical Peano curve in $L_\infty$ metric, where $m$ is an evolvent construction parameter. Problems of numerical construction of the evolvents and the corresponding theory are considered in detail in \cite{Strongin2000}.

By using this kind of mapping it is possible to reduce the multidimensional problem~(\ref{problem}) to a univariate problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]

The considered dimensionality reduction scheme juxtaposes a
multidimensional problem with lipschitzian functions to a univariate
problem where the corresponding functions satisfy the uniform H{\"o}lder
condition (see \cite{Strongin2000}), i.e.
\[
\left|g_i(y(x'))-g_i (y(x''))\right| \leq H_i \left|x'-x'' \right|^{1/N}, \; x',x''\in [0,1], \; 1\leq i \leq m+1.
\]
Here $N$ is the dimensionality of the initial multidimensional problem and
the coefficients $H_i$ are related with the Lipschitz constants $L_i$ of
the initial problem by the inequalities $H_i \leq 2L_i \sqrt{N+3}$.

\subsection{Nested optimization scheme}

Nested optimization scheme is based on relation (see \cite{Grishagin2001})
\begin{equation}\label{nested}
\min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}...\min_{u_M\in D_M }{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
\end{equation}
which allows replacing the solving of multidimensional problem (\ref{problem}) by solving a family of subproblems related to each other recursively.
Here we consider vector $y$ as a vector of block variables
\begin{equation}
y=(y_1,y_2,...,y_N)=(u_1,u_2,...,u_M),
\end{equation}
where the $i$-th block variable $u_i$ is a vector of $N_i$ components of vector $y$, taken serially i.e. $u_1=(y_1,y_2,...,y_{N_1})$, $u_2=(y_{N_1+1},y_{N_1+2},...,y_{N_1+N_2})$,..., $u_M=(y_{N-N_M+1},y_{N-N_M+2},...,y_{N})$, at that $N_1+N_2+...+N_M=N$. 
The subdomains $D_i, 1 \leq i \leq M$, are projections of initial search domain $D$ onto the subspaces corresponding to the variables $u_i, 1 \leq i \leq M$.

\Russian 
¬ проведенном исследовании данна€ схема примен€лась при M=2, т.е. использовалс€ только один уровень вложенности
\begin{equation}
\min_{y \in D}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}= \min_{u_1\in D_1}\min_{u_2\in D_2}{\left\{\varphi(y): \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}.
\end{equation}

\section{Organization of Parallel Computing}\label{sec:4}

\section{Numerical experiments}\label{sec:5}

\begin{itemize}
	\item one could control the size of the feasible domain $Q_{m+1}$ with respect to the whole domain $D$;
	\item the global minimizer of the objective function is known a priori taking into account the constraints;
	\item the global minimizer of the objective function without accounting for the constraints is out of the feasible domain $Q_{m+1}$  (with the purpose of simulating the behavior of the constraints and the objective function in the applied constrained optimization problems).
\end{itemize}

\begin{figure}[ht]
\begin{minipage}[h]{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{vag_06.png} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}[h]{0.5\linewidth}
\center{\includegraphics[width=1.0\linewidth]{GKLS_04.png} \\ (b)}
\end{minipage}
\center{{\bf Figure 1.} The level lines of (a) hard and (b) simple subproblems }
\end{figure}

The numerical experiments were carried out using two classes of problems (\textit{Simple} ans \textit{Hard}) with $N=5$. The problem was considered to be solved if the algorithm generates a trial point $y^k$ in $\delta$-vicinity of the global minimizer, i.e. $\left\|y^k-y^\ast\right\|\leq\delta$ . The size of the vicinity was selected as $\delta=0.01\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain $D$. The maximum allowable number of iterations was $K_{max} = 10^6$.

\Russian
ѕервый эксперимент проведем при решении серий задач Simple и Hard на одном узле, полностью задействовать оба доступных CPU (т.е. p=16 физических €дер). ¬ таблице 1 приведено среднее врем€ (в секундах), потребовавшихс€ дл€ решени€ задач серии.

\begin{table}[ht]
\center{{\bf Table 1.} Average time for solving the problem on one node}
\begin{center}
\begin{tabular}{cc}
\hline
Problems & Time, sec. \\
\hline
\textit{Hard} & 52  \\
\textit{Simple} & 51 \\
\hline
\end{tabular}
\end{center}
\end{table}

\Russian
—ледующий эксперимент проведем, задействовав p узлов по три графических ускорител€ на каждом из них. “ак, при p=32 было задействовано 94 GPU ускорител€, на каждом из которых Ц 2688 CUDA-€дер.  ¬ таблице 2 приведено врем€ работы алгоритма, а в таблице 3 Ц ускорение по отношению к запуску на одном узле.

\begin{table}[ht]
\center{{\bf Table 2.} Average time for solving the problem on $p$ nodes}
\begin{center}
\begin{tabular}{cccc}
\hline
Problems & $p = 8$ & $p=16$ & $p=32$ \\
\hline
\textit{Hard} & 11.51 & 5.53 & 0.54 \\
\textit{Simple} & 2.04 & 1.50 & 0.47\\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[ht]
\center{{\bf Table 3.} Speedup with respect to one node}
\begin{center}
\begin{tabular}{cccc}
\hline
Problems & $p = 8$ & $p=16$ & $p=32$ \\
\hline
\textit{Hard} & 5 & 9 & 96\\
\textit{Simple} & 25 & 34& 109\\
\hline
\end{tabular}
\end{center}
\end{table}

Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhni Novgorod. The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 Gb RAM, and three NVIDIA Kepler K20’ GPUs (2688 CUDA cores, 6 Gb GDDR5).

\section{Conclusion}
